---
title: 'Lecture 6: Two Means'
author: "Dr Jennifer Mankin"
date: "25/01/2020"
output: html_document
---

# Overview: Two Means

* Study design
  + Defining variables
  + What question are we asking?
* Statistical analysis
  + Student's t-test
  + Under the hood: drawing lines
* Two means in R
  + Welcome to `lm()`!
  + Getting and interpreting results

## Study design

* 

## Statistical analysis



## Two means in R

* Welcome to `lm()`!
* Getting and interpreting results

### Welcome to `lm()`!

* Today's new function is a very important one
  + We will use it for the rest of the term and...
  + It will be very important next year as well!
* What does it do?
  + Creates a **l**inear **m**odel -> `lm()`
  + A statistical model that looks like a line
* How does the linear model work?
  + Much more next week!

### Basic Anatomy of `lm()`

* `lm(outcome ~ predictor, data = data)`
* Translation: 
  + Create a linear model modelling the relationship between
  + An outcome and a predictor
  + Using the data in `data`
* **Try it**: write down the command to predict `attract` from `colour` in our `elliot` data

### Anatomy of `lm()`

`lm(attract ~ colour, data = elliot)`

* Create a linear model that models
* Attractiveness score based on colour group
* Using the data `elliot`

### What will this model actually do?

`attract ~ colour`

* `attract` is a **continuous** variable: each participant's mean score on three rating scales of attractiveness
* `colour` is a **categorical** variable: whether each participant saw the `red` condition or the `white` condition
  + Independent: each participant saw only one of the two

### What will this model actually do?

`attract ~ colour`

1. Group participants by `colour`
2. Calculate the mean `attract` score for each `colour` group
3. Create a linear model that captures the **difference in those means**

## Let's have a look!

```{r}
attract_w <- rnorm(2000, mean = 4, sd = 0.5)
attract_r <- rnorm(2000, mean = 7, sd = 0.5)
attract <- c(attract_w,attract_r)
elliot <- tibble(
  colour = factor(rep(c("white", "red"), each = 2000)),
  attract = attract
)

ggplot(elliot, aes(attract, fill = colour)) + 
   geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity')

# Creating point plot with CIs
ggplot(elliot, aes(x = colour, y = attract, fill = colour)) + 
  stat_summary(fun.data = "mean_cl_normal") # doesn't work atm but might at work!

```


### `attract ~ colour`

* Are these means **different**?
  + Yes: the `colour` groups have different means of `attract`
* Are these means different **enough**?
  + How big of a difference is big enough to be meaningful?
  + How big of a difference is big enough to be **significant**?

### `attract ~ colour`

Our model will give us two important things:
* An **estimate** of the relationship between `colour` and `attract`
  + Here: the difference between the means in the white and red groups
* The **probability** of finding a difference in means of this size (or greater) occuring if seeing a white or red border has no effect on attractiveness
  + i.e., if the null hypothesis is true

### Getting results

Let's run the model and see what it is telling us

```{r}
elliot_m <- lm(attract ~ colour, data = elliot)
summary(elliot_m)
```


